Ch. 2 Dissertation draft

### 2020-02-12 notebooks/article_draft.Rmd
Adds chunks to download REMA DEMs and calculate/extract slope and aspect

# Make df of relevant data in long format
accum.long = data %>% select(Lat, Lon, elev, Year) %>% 
  mutate(accum = mode.accum, sd = sqrt(var.accum))

@@

# Recombine accum data in long format with location data and rearrange
cores.accum = left_join(core_long, core.locs, by = "Site")  %>% arrange(Site, Year) %>% 
  select(Site, Latitude, Longitude, Elevation, Year, Accum) %>% 
  mutate(Year = as.integer(Year))

@@

@@

# Extract relevant model data and store in nested tibble
accum.loc_data = select(accum.loc_PS, -data, -accum_mu) %>% 
  mutate(data = map(accum.lm_loc, 
                    ~ select(augment(.), Year, accum, .fitted, .se.fit, .resid, .std.resid)))

# Extract model parameter estimates and p values and store as location-based tibble
tmp.tidy = map(accum.lm_loc, ~ tidy(.x))
trend.loc = accum.loc_PS %>% 
  select(Easting, Northing, elev, accum_mu) %>% 
  mutate(intrcpt = simplify(map(tmp.tidy, "estimate") %>% map(1))) %>%
  mutate(coeff_yr = simplify(map(tmp.tidy, "estimate") %>% map(2))) %>% 
  mutate(std.err = simplify(map(tmp.tidy, "std.error") %>% map(2))) %>% 
  mutate(p.val_yr = simplify(map(tmp.tidy, "p.value") %>% map(2)))

@@

# Extract relevant model data and store in nested tibble
cores.loc_data = select(cores.loc_PS, -Data, -accum_mu, -data.length) %>% 
  mutate(data = map(cores.lm_loc, 
                    ~ select(augment(.), Year, Accum, .fitted, .se.fit, .resid, .std.resid)))

# Extract model parameter estimates and p values and store as location-based tibble
tmp.tidy = map(cores.lm_loc, ~ tidy(.x))
cores.trend = cores.loc_PS %>% 
  select(Site, Easting, Northing, Elevation, accum_mu) %>% 
  mutate(intrcpt = simplify(map(tmp.tidy, "estimate") %>% map(1))) %>%
  mutate(coeff_yr = simplify(map(tmp.tidy, "estimate") %>% map(2))) %>% 
  mutate(std.err = simplify(map(tmp.tidy, "std.error") %>% map(2))) %>% 
  mutate(p.val_yr = simplify(map(tmp.tidy, "p.value") %>% map(2)))

@@

test.data = select(test.smp, -data, -accum_mu) %>% 
  mutate(data = map(test.lm, 
                    ~ select(augment(.), Year, accum, .fitted, .se.fit, .resid, .std.resid)))

@@

```{r}
# Find all mosaics that intersect the accumulation dataset
index.file = "/media/durbank/WARP/Research/Antarctica/Data/DEMs/REMA/REMA_Tile_Index_Rel1.1/REMA_Tile_Index_Rel1.1.shp"
dem.index = st_read(index.file)
ggplot() + geom_sf(data = Ant.outline, color = 'grey') + geom_sf(data = dem.index)
keep.idx = unique(unlist(st_intersects(accum.PS, dem.index)))
mosaic.download = dem.index[keep.idx,] %>% mutate(fileurl = as.character(fileurl))

# output.dir = file.path("/media/durbank/WARP/Research/Antarctica/Data/DEMs/REMA/WAIS_divide")
output.dir = tk_choose.dir(caption = "Select directory in which to save REMA mosaic tiles")

# Download files from REMA ftp server
for (file in 1:nrow(mosaic.download)) {
  tmp.file = tempfile()
  dir.name = basename(dirname(mosaic.download$fileurl[file]))
  # dir.create(file.path(output.dir, dir.name))

  # If not already downloaded, download tile and extract
  if (!dir.exists(file.path(output.dir, dir.name))) {
    download.file(mosaic.download$fileurl[file], tmp.file)
    untar(tmp.file, exdir = file.path(output.dir, dir.name))
  }
  unlink(tmp.file)
}
```

```{r}
# tic("Raster analysis")
# 
# dem.file = "/media/durbank/WARP/Research/Antarctica/Data/DEMs/REMA/WAIS_divide/r.slope.tif"
# dem.slope = raster(dem.file)
# accum.loc_PS = accum.loc_PS %>% mutate(slope = extract(dem.slope, accum.PS))
# 
# toc()
```

>>>>

# Make df of relevant data in long format
accum.long = data %>% dplyr::select(Lat, Lon, elev, Year) %>% 
  mutate(accum = mode.accum, sd = sqrt(var.accum))

@@

# Recombine accum data in long format with location data and rearrange
cores.accum = left_join(core_long, core.locs, by = "Site")  %>% arrange(Site, Year) %>% 
  dplyr::select(Site, Latitude, Longitude, Elevation, Year, Accum) %>% 
  mutate(Year = as.integer(Year))

@@

```{r}
## This code chunk serves to find which REMA DEM 8-meter tiles intersect our dataset 
## and downloads any tiles not already present in the selected directory "output.dir"

# REMA.dir = tk_choose.dir(caption = "Select directory in which to save REMA mosaic tiles")
REMA.dir = file.path("G:/Research/Antarctica/Data/DEMs/REMA/tiles_8m_v1.1")
# REMA.dir = file.path("/media/durbank/WARP/Research/Antarctica/Data/DEMs/REMA/tiles_8m_v1.1")

# Find all mosaics that intersect the accumulation dataset
index.file = file.path(dirname(REMA.dir), "REMA_Tile_Index_Rel1.1/REMA_Tile_Index_Rel1.1.shp")
dem.index = st_read(index.file)
# ggplot() + geom_sf(data = Ant.outline, color = 'grey') + geom_sf(data = dem.index)
keep.idx = unique(unlist(st_intersects(accum.PS, dem.index)))
mosaic.download = dem.index[keep.idx,] %>% mutate(fileurl = as.character(fileurl))

# Download files from REMA ftp server
for (file in 1:nrow(mosaic.download)) {
  tmp.file = tempfile()
  dir.name = basename(dirname(mosaic.download$fileurl[file]))

  # If not already downloaded, download tile and extract
  if (!dir.exists(file.path(REMA.dir, dir.name))) {
    download.file(mosaic.download$fileurl[file], tmp.file)
    untar(tmp.file, exdir = file.path(REMA.dir, dir.name))
  }
  unlink(tmp.file)
}
```

```{r}
tic("Raster extraction")

# Function to find which accum points are within a given REMA 8-meter mosaic tile 
# and extract elevation, slope, and aspect from the REMA raster data
get_topo_vals = function(tile, accum.pts, tile.dir) {

  # Logical to filter accum data to those members within the tile extent
  pts.log = st_intersects(accum.pts, tile, sparse = FALSE)[,1]
  accum.i = accum.pts %>% filter(pts.log)

  # Generate file paths to DEM, slope raster and aspect raster
  dem.path = file.path(tile.dir, as.character(tile$tile), 
                        paste(as.character(tile$tile), "_8m_dem.tif", sep = ""))
  slope.path = file.path(tile.dir, as.character(tile$tile), 
                        paste(as.character(tile$tile), "_8m_slope.tif", sep = ""))
  aspect.path = file.path(tile.dir, as.character(tile$tile), 
                        paste(as.character(tile$tile), "_8m_aspect.tif", sep = ""))

  # Check if slope/aspect rasters exist. If they don't, calculate them from DEM 
  # and save to REMA directory
  if (!file.exists(slope.path)) {
    tile.dem = raster(dem.path)
    tile.slope = terrain(tile.dem, opt='slope', unit='degrees')
    writeRaster(tile.slope, filename = slope.path, format = "GTiff")
    tile.aspect = terrain(tile.dem, opt='aspect', unit='degrees')
    writeRaster(tile.aspect, filename = aspect.path, format = "GTiff")
  }

  # Generate raster stack of DEM, slope, and aspect 
  # and extract raster values at accum points
  tile.stack = stack(list(dem.path, slope.path, aspect.path))
  stack.vals = extract(tile.stack, accum.i)

  # Get index values of accum data subset within the global accum dataset
  pts.idx = (1:length(pts.log))[pts.log]

  # Return index positions and raster stack values as list object
  return(list(pts.idx, stack.vals))
}

# Add variables for REMA elevation, slope, and aspect
accum.PS = accum.PS %>% mutate(elev.REMA = NA, slope = NA, aspect = NA)

for (i in 1:nrow(mosaic.download)) {

  # Get current tile with coverage in dataset
  tile = mosaic.download[i,]

  # Calculate and extract elevation, slope, and aspect
  vals.list = get_topo_vals(tile, accum.PS, REMA.dir)
  topo.idx = vals.list[[1]]
  topo.vals = vals.list[[2]]

  # Assign topo values to accum tbl
  accum.PS$elev.REMA[topo.idx] = topo.vals[,1]
  accum.PS$slope[topo.idx] = topo.vals[,2]
  accum.PS$aspect[topo.idx] = topo.vals[,3]
}

@@

# Extract relevant model data and store in nested tibble
accum.loc_data = dplyr::select(accum.loc_PS, -data, -accum_mu) %>% 
  mutate(data = map(accum.lm_loc, 
                    ~ dplyr::select(augment(.), Year, accum, .fitted, .se.fit, .resid, .std.resid)))

# Extract model parameter estimates and p values and store as location-based tibble
tmp.tidy = map(accum.lm_loc, ~ tidy(.x))
trend.loc = accum.loc_PS %>% 
  dplyr::select(Easting, Northing, elev, accum_mu) %>% 
  mutate(intrcpt = simplify(map(tmp.tidy, "estimate") %>% map(1))) %>%
  mutate(coeff_yr = simplify(map(tmp.tidy, "estimate") %>% map(2))) %>% 
  mutate(std.err = simplify(map(tmp.tidy, "std.error") %>% map(2))) %>% 
  mutate(p.val_yr = simplify(map(tmp.tidy, "p.value") %>% map(2)))

@@

# Extract relevant model data and store in nested tibble
cores.loc_data = dplyr::select(cores.loc_PS, -Data, -accum_mu, -data.length) %>% 
  mutate(data = map(cores.lm_loc, 
                    ~ dplyr::select(augment(.), Year, Accum, .fitted, .se.fit, .resid, .std.resid)))

# Extract model parameter estimates and p values and store as location-based tibble
tmp.tidy = map(cores.lm_loc, ~ tidy(.x))
cores.trend = cores.loc_PS %>% 
  dplyr::select(Site, Easting, Northing, Elevation, accum_mu) %>% 
  mutate(intrcpt = simplify(map(tmp.tidy, "estimate") %>% map(1))) %>%
  mutate(coeff_yr = simplify(map(tmp.tidy, "estimate") %>% map(2))) %>% 
  mutate(std.err = simplify(map(tmp.tidy, "std.error") %>% map(2))) %>% 
  mutate(p.val_yr = simplify(map(tmp.tidy, "p.value") %>% map(2)))

@@

test.data = dplyr::select(test.smp, -data, -accum_mu) %>% 
  mutate(data = map(test.lm, 
                    ~ dplyr::select(augment(.), Year, accum, .fitted, .se.fit, .resid, .std.resid)))

@@
